---
title: "Cleaning Berries"
author: "MA615"
date: "10/6/2020"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
library(knitr)
library(tidyverse)
library(magrittr)
library(kableExtra)

opts_chunk$set(echo = FALSE, 
               warning = FALSE,
               message = FALSE)
```

##  Acquire and read the data

These data were collected from the USDA database selector: <a href="https://quickstats.nass.usda.gov">https://quickstats.nass.usda.gov</a>

The data were <a href="https://quickstats.nass.usda.gov/results/D416E96E-3D5C-324C-9334-1D38DF88FFF1">stored online</a> and then downloaded as a CSV file.


```{r}

## read the data

ag_data <- read_csv("berries.csv", col_names = TRUE)

## look at number of unique values in each column
ag_data %>% summarize_all(n_distinct) -> aa


## make a list of the columns with only one unique value
bb <- which(aa[1,]==1)

## list the 1-unique valu column names 
cn <- colnames(ag_data)[bb]

```

Data selected from the NASS database often has columns without any data or with a single repeated Values.  The berries data had only 8 out of 21 columns containing meaningful data.

```{r}


## remove the 1-unique columns from the dataset
ag_data %<>% select(-all_of(bb))

aa %<>% select(-all_of(bb)) 


## State name and the State ANSI code are (sort of) redundant
## Just keep the name
ag_data %<>% select(-4)
aa %<>% select(-4) 


kable(head(ag_data)) %>%
  kable_styling(font_size=12)


```

<hr>
<br>



```{r}
berry <- unique(ag_data$Commodity)
nberry <- length(berry)

```

This table contains informaton about `r nberry` berries: blueberries, raspberries, and strawberries.

When the data have been cleaned and organized, the three kinds of berries will be separted into tables with the same stucture so that they can be compared.  So, working with Blueberries along demonstrates how the data will be cleaned and organized for all three kinds of berries. Only the "YEAR" time periond will be considered.

## Blueberries

```{r}
bberry <- ag_data %>% filter((Commodity=="BLUEBERRIES") & (Period=="YEAR"))
bberry %<>% select(-c(Period, Commodity))   

#### Does every Data Item begin with "
sum(str_detect(bberry$`Data Item`, "^BLUEBERRIES, ")) == length(bberry$`Data Item`)


# di <- bberry$`Data Item`
# di_m <- str_split(di, ",", simplify=TRUE)
# dim(di_m)
# 
# unique(di_m[,1])
# di_m <- di_m[,2:4]

bberry %<>% separate(`Data Item`, c("B","type", "meas", "what"), sep = ",") 
bberry %<>% select(-B)

# head(bberry$type, n=20)
# ty <- str_split(bberry$type, " ", simplify=TRUE)
# head(ty, n=20)

bberry %<>% separate(type,c("b1", "type", "b2", "lab1", "lab2"), " ")

bberry %<>% select(-c(b1,b2)) 

bberry[is.na(bberry)] <- " "  ## OK now Data Item has been split into parts

## onto Domain

```


```{r}


# bberry$Domain %>% unique()

bberry %<>% separate(Domain, c("D_left", "D_right"), sep = ", ")

# bberry$D_left %>% unique()
# bberry$D_right %>% unique()

bberry[is.na(bberry)] <- " "

## And now Domain Category


## bberry$`Domain Category` %>% unique()

bberry %<>% separate(`Domain Category`, c("DC_left", "DC_right"), sep = ", ")

## looks like DC_left combines labels

head(bberry$DC_left %>% unique(),n=20)
head(bberry$DC_right %>% unique(), n=20)


## work on DC_left first

bberry %<>% separate(DC_left, c("DC_left_l", "DC_left_r"), sep = ": ")

## bberry$DC_left_l %>% unique()
## bberry$DC_left_r %>% unique()

## now work on DC_right

head(bberry$DC_right %>% unique(), n=20)

bberry %<>% separate(DC_right, c("DC_right_l", "DC_right_r"), sep = ": ") 


bberry[is.na(bberry)] <- " "

##  OK now we need to eliminate the redundancy
```



```{r}

## fine and remove redundant columns
 
## paste(bberry$D_left,bberry$DC_left_l) %>% unique
## returns -- "TOTAL NOT SPECIFIED"   "CHEMICAL CHEMICAL"     "FERTILIZER FERTILIZER"

## remove column bberry$DC_left_l


bberry %<>%  select(-DC_left_l) 

## test

# sum(bberry$D_right == bberry$DC_right_l)
# [1] 7419
# > bberry$DC_left_r %>% unique()
# [1] " "           "(NITROGEN)"  "(PHOSPHATE)" "(POTASH)"    "(SULFUR)"  

## remove column DC_right_l

bberry %<>% select(-DC_right_l)

## Test for lab1, lab2

# paste(bberry$lab1, bberry$lab2) %>% unique()
# [1] "ACRES HARVESTED" "PRODUCTION  "    "YIELD  "         "   "             "APPLICATIONS  " 
# [6] "TREATED  " 


bberry %<>% mutate(label = paste(lab1,lab2)) 

## test for necisity of "chemical" in col D_left


# paste(bberry$D_left, bberry$D_right) %>% unique()
# [1] "TOTAL  "              "CHEMICAL FUNGICIDE"   "CHEMICAL HERBICIDE"   "CHEMICAL INSECTICIDE"
# [5] "CHEMICAL OTHER"       "FERTILIZER  "   


## remove "Chemical" and joint the columns

bberry %<>% mutate(D_left = "CHEMICAL", D_left = "") 



bberry %<>% mutate(Chemical=paste(D_left, D_right)) 

bberry %<>% select(-c(D_left, D_right)) 


bberry %<>% select(Year, State, type, what, meas, label, DC_left_r, DC_right_r, Chemical, Value )

```


```{r}

###  Now the problem is that we have entries in both the "what" and "meas" columns
##  that begin  "MEASURED IN"
##  how many are there

## in the column "what"
cnt_1 <- str_detect(bberry$what, "MEASURED IN")
sum(cnt_1)

## in the column "meas"

cnt_2 <- str_detect(bberry$meas, "MEASURED IN")
sum(cnt_2)

## We want to put them all in the same column
## So, we will separate them from their current column and put them into
## two columns -- then we will test to make sure there aren't any overlaps
## and then merge the two columns

## we're going to use PURRR.  We need a simple function that takes a logical
## variable and a second variable.  It returns the second variable if the logical
## variable is true and returns a blank if it is false


f1 <- function(a,b){
  if(a){
    return(b)
  }else{
      return("")
  }
}
#################################################
##  try it

f1(TRUE,"hi")

f1(!TRUE, "hi")

## now try it with map2()

f1_log <- c(FALSE, TRUE, TRUE)
f1_str <- c("one", "two", "three")

map2(f1_log, f1_str, f1)

#########################################################
## now let's separate the "MEASURED IN" entries in the meas column
## form an index of the entries to be separated out

index_meas <- str_detect(bberry$meas, "MEASURED IN")

## verify the first six values against the dats bberry
head(index_meas)
new <- map2(index_meas, bberry$meas, f1)
new <- unlist(new)
head(new, n=20)
######################################################

bberry %<>% mutate(m_in_1 = unlist(map2(index_meas, bberry$meas, f1))) 

# 
#
# b_f <- bberry
# 
#  
#  
#  l1 <- "MEASURED IN LB"
#  l2 <- "howdy"
#  l3 <- "MEASURED IN LB / ACRE"
# 
#  str_replace(c(l1,l2,l3), "MEASURED IN.*$", "")
# 
# tmp <- str_replace(l4, "MEASURED IN.*$", "")
# 
# b_f %<>%  mutate(meas = str_replace(b_f$meas, "MEASURED IN.*$", "" ))
# 
# Check on whether it worked
#
# cnt_l <- str_detect(b_f$meas, "MEASURED IN")
# sum(cnt_l)
#

bberry %<>% mutate(meas = str_replace(bberry$meas, "MEASURED IN.*$", ""))

## Check
cnt_3 <- str_detect(bberry$meas, "MEASURED IN")
sum(cnt_3)


#########################
## Now we will do the same thing with the 
## "what" column  

### index of cells to be isolated
index_what <- str_detect(bberry$what, "MEASURED IN")
sum(index_what)

### create a column of the isolated cells
bberry %<>% mutate(m_in_2 = unlist(map2(index_what, bberry$what, f1))) 

###  eliminate the isolated cells from the original column
bberry %<>% mutate(what = str_replace(bberry$what, "MEASURED IN.*$", ""))

### test that theere are no more "MEASURED IN" cells in the original column
cnt_what <- str_detect(bberry$what, "MEASURED IN")
sum(cnt_what)

### Check for overlaps


bberry %<>% mutate(units = str_trim(paste(m_in_1, m_in_2))) 

bberry$units %>% unique()


```




```{r}

## now let's clean it up 

bberry$what %>% unique()  ## rename Avg

bberry$meas %>% unique()  ## rename marketing

bberry$label %>% unique() ## rename harvest 

bberry$DC_left_r %>% unique() # rename chemical_family

tmp <- bberry$DC_right_r %>% unique() # rename materials --213

tmp <- bberry$Value %>% unique() # values

tmp <- bberry$units %>% unique() # Measures




bberry %<>% rename(Avg = what)
bberry %<>% rename(Marketing = meas, Harvest = label, Chem_family = DC_left_r, Materials = DC_right_r, Measures = units)

colnames(bberry)

bberry %<>% select(Year, State, type, Marketing, 
                   Measures, Avg, Harvest, Chem_family,
                   Materials, Chemical, Value )

str_trim(paste(bberry$Marketing, bberry$Harvest)) %>% unique

###  these belong in one column

bberry %<>% mutate(production = str_trim(paste(Marketing, Harvest)))

bberry %<>% select(Year, State, type, production, Measures,
                   Avg, Chem_family, Materials, Chemical, Value)


## I missed this one !!

bberry %<>% mutate(Chemical = str_trim(paste(Chem_family, Chemical)))

bberry %<>% select(Year, State, type, production, Avg, Measures, Materials, Chemical, Value)

```


```{r}

### let's look at chemicals being applied to food

unfood <- bberry %<>% filter(production=="APPLICATIONS") 


unfood %<>% filter(Value != "(D)")

unfood %<>% filter(Value !=  "(NA)")

unfood %<>% filter(Measures == "MEASURED IN LB / ACRE / APPLICATION")

unfood$type %>% unique()                  ## BEARING
unfood$production %>% unique()            ## APPLICATIONS
unfood$Avg %>% unique()                   ## AVG
unfood$Measures %>% unique()              ## MEASURED IN LB / ACRE / APPLICATION
unfood$Materials %>% unique()      ## long list!!        
unfood$Chemical %>% unique()


unfood_1 <- unfood %>%  select(Year, State, Chemical, Value)

unfood_1 %<>% pivot_wider(names_from = Chemical, values_from = Value)

a <- unfood_1$FUNGICIDE



#kable(head(bberry, n=10)) %>% kable_styling(font_size=12)

```



### Units



